\textsc{HubbleCLIP}: Associating Astronomical Observations and Natural Language with Multi-Modal Models

### 2024-01-03 paper/main.tex
Paper update

CLIP \citep[Contrastive Language-Image Pretraining;][]{radford2021learning} is a multi-modal model pre-trained on a large corpus of image-text pairs via weak supervision using a contrastive loss. Given a minibatch $\mathcal{B}$ of $|\mathcal{B}|$ image-text pairs $\{(I_i, T_i)\}$, the goal is to align the learned representations of corresponding (positive) pairs $(I_i, T_i)$ while repelling the representations of unaligned (negative) pairs $(I_i, T_{j\neq i})$. Image and text encoders $f(\cdot)$ and $g(\cdot)$ are used to map images and text to a common embedding space. The standard softmax-based bidirectional variation of the InfoNCE~\citep{oord2018representation} contrastive loss function, as used by CLIP, is particularly effective for multimodal learning. This bidirectionality is crucial as it ensures the model learns to map both images to text and text to images with equal importance. This symmetry in learning is essential for tasks that require a mutual understanding and interchangeability between visual and textual representations, such as image captioning, text-to-image synthesis, and cross-modal retrieval. The bidirectional loss is given by \citep{radford2021learning}
\begin{equation}
\mathcal L(\mathcal B) = -\frac{1}{2|\mathcal{B}|} \sum_{i=1}^{|\mathcal{B}|}\left(\overbrace{\log \frac{e^{{x}_i \cdot {y}_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{{x}_i \cdot {y}_j / \tau}}}^{\text {image } \rightarrow \text { text softmax }}+\overbrace{\log \frac{e^{{x}_i \cdot {y}_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{{x}_j \cdot {y}_i / \tau}}}^{\text {text } \rightarrow \text { image softmax }}\right)
\end{equation}
where ${x}_i=\frac{f\left(I_i\right)}{\left\|f\left(I_i\right)\right\|_2}$ and ${y}_i=\frac{g\left(T_i\right)}{\left\|g\left(T_i\right)\right\|_2}$ are the normalized representations of the $i$-th image and text, respectively, and $\tau$ is a (typically learnable) temperature hyperparameter. 
% This bidirectional approach not only enhances the model's ability to create robust and versatile embeddings but also aligns with the intuitive understanding of how humans interact with and interpret multimodal information, further bridging the gap between artificial and natural intelligence in multimodal contexts.

@@

Models were instantiated and trained using the \package{Jax} \citep{jax2018github} ecosystem. The model is trained over 50,000 steps with 5000 linear warmup steps and cosine decay using the AdamW optimizer \citep{DBLP:conf/iclr/LoshchilovH19,DBLP:journals/corr/KingmaB14} with peak learning rate of either $10^{-5}$ or $10^{-6}$ and weight decay $10^{-3}$. Training takes approximately 6 hours on 4 Nvidia A100 GPUs.

>>>>

CLIP \citep[Contrastive Language-Image Pretraining;][]{radford2021learning} is a multi-modal model pre-trained on a large corpus of image-text pairs via weak supervision using a contrastive loss. Given a minibatch $\mathcal{B}$ of $|\mathcal{B}|$ image-text pairs $\{(I_i, T_i)\}$, the goal is to align the learned representations of corresponding (positive) pairs $(I_i, T_i)$ while repelling the representations of unaligned (negative) pairs $(I_i, T_{j\neq i})$. Image and text encoders $f(\cdot)$ and $g(\cdot)$ are used to map images and text to a common embedding space. The standard softmax-based bidirectional variation of the InfoNCE~\citep{oord2018representation} contrastive loss function, as used by CLIP, is particularly effective for multimodal learning. This bidirectionality is crucial as it ensures the model learns to map both images to text and text to images with equal importance. This symmetry in learning is essential for tasks that require a mutual understanding and interchangeability between visual and textual representations, such as image captioning, text-to-image synthesis, and cross-modal retrieval. The bidirectional loss is given by \citep{radford2021learning}
\begin{equation}
\mathcal L(\mathcal B) = -\frac{1}{2|\mathcal{B}|} \sum_{i=1}^{|\mathcal{B}|}\left(\underbrace{\log \frac{e^{{x}_i \cdot {y}_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{{x}_i \cdot {y}_j / \tau}}}_{\text {image } \rightarrow \text { text softmax }}+\underbrace{\log \frac{e^{{x}_i \cdot {y}_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{{x}_j \cdot {y}_i / \tau}}}_{\text {text } \rightarrow \text { image softmax }}\right)
\end{equation}
where ${x}_i=\frac{f\left(I_i\right)}{\left\|f\left(I_i\right)\right\|_2}$ and ${y}_i=\frac{g\left(T_i\right)}{\left\|g\left(T_i\right)\right\|_2}$ are the normalized representations of the $i$-th image and text, respectively, and $\tau$ is a learnable temperature hyperparameter. 
% This bidirectional approach not only enhances the model's ability to create robust and versatile embeddings but also aligns with the intuitive understanding of how humans interact with and interpret multimodal information, further bridging the gap between artificial and natural intelligence in multimodal contexts.

@@

Models were instantiated and trained using the \package{Jax} \citep{jax2018github} ecosystem. The model is trained over 20,000 steps with 5000 linear warmup steps and cosine decay using the AdamW optimizer \citep{DBLP:conf/iclr/LoshchilovH19,DBLP:journals/corr/KingmaB14} with peak learning rate of $10^{-5}$ and weight decay $10^{-3}$. Training takes approximately 3 hours on 4 Nvidia A100 GPUs.

% Describe evals
